---
title: "Abiodun_Abioye_0778983_Final Project Write Up"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.width = 7,fig.height = 5)
# Load useful packages
library(tidyverse)
library(hablar)
library(gridExtra)
library(ggridges)
library(plotly)
library(here)
library(modelr)
library(patchwork)
library(webshot)
options(na.action = na.warn)
```

```{r echo=FALSE,message=FALSE}
# Load Dataset
CreditCard <- read_csv("Data/CreditCard.csv")
```

```{r echo=FALSE,message=FALSE}
# Convert variables to correct datatypes
CreditCard_df <- convert(CreditCard, int(age, dependents, months, active))
```

### Introduction
This document shows a detailed analysis on the CreditCard dataset. Answers to analysis questions are provided. 

```{r echo=FALSE,message=FALSE}
#Take out outlier values (i.e. Age < 18)
new_CreditCard <- CreditCard_df %>%
  filter(age >= 18 )
```

##### Weak, positive linear relationship between age and income. 

Older people tend to make more money. Self employment status (no and yes) is evenly distributed across age and income bracket. Regular workers (who are not self employed) have a more positive relationship. This means that the yearly income of regular workers increases more with age and self-employed people.

```{r echo=FALSE}
ggplot(new_CreditCard, aes(x=income, y=age, colour = selfemp)) + geom_point()+ geom_smooth(method = 'lm', se=FALSE) + labs(x= "Yearly income(USD10,000)", y="Age")
```


##### A 1 year increase in age leads to an increase of \$500 in yearly income

```{r echo=FALSE}
m3 <- lm(income ~ age, data = new_CreditCard)
summary(m3)
```

##### Correlation between income and age. 

```{r echo=FALSE}
cor(new_CreditCard$income, new_CreditCard$age)
```

##### No clear relationship between age and expenditure. 
```{r echo=FALSE}
ggplot(new_CreditCard, aes(x=expenditure, y=age, colour = selfemp)) + geom_point()+ geom_smooth(method = 'lm', se=FALSE) + labs(x= "Monthly Credit Card Expenditure", y="Age")
```

##### Correletation coefficient - Age and Expenditure
```{r echo=FALSE}
cor(new_CreditCard$age, new_CreditCard$expenditure)
```

##### Log Transformation of expenditure. 

```{r echo=FALSE}
p1 <- ggplot(x = expenditure, data = new_CreditCard)
p1 <- qplot(x = expenditure, data = new_CreditCard)
p2 <- qplot(x = log10(income + 1), data = new_CreditCard)
grid.arrange(p1, p2, ncol = 1)
```

##### For every \$1,000 in income, there will be an increase in average monthly credit card expenditure of \$45.240. 

```{r echo=FALSE}
m1 <- lm(expenditure ~ income, data = new_CreditCard)
summary(m1)
```

##### Very Weak, positive linear relationship between Yearly income (x axis) and average monthly credit card expenditure (y axis). 

```{r echo=FALSE}
fig1 <- plot_ly(new_CreditCard, x = ~income, y = ~expenditure, type  = 'scatter', mode = 'markers') 
                
fig <- subplot(fig1)
fig
```

##### Correlation coefficient between Income and Expenditure. 

```{r echo=FALSE}
cor(new_CreditCard$income, new_CreditCard$expenditure)
```

```{r echo=FALSE}
ggplot(new_CreditCard, aes(x=income,y=expenditure )) + geom_point()+ geom_smooth(method="lm", colour="red")
```

##### Add & Plot Predictions

```{r echo=FALSE}
grid <- new_CreditCard %>%
  data_grid(income)
grid <-grid %>%
  add_predictions(m1)

ggplot(new_CreditCard, aes(income)) + geom_point(aes(y = expenditure)) + geom_line(aes(y = pred), data = grid, colour = "red", size = 1)
new_CreditCard <- new_CreditCard %>%
  add_residuals(m1)
```

##### Residuals Plot

```{r echo=FALSE}
a <- ggplot(new_CreditCard, aes(resid,expenditure))+geom_ref_line(h = 0)+geom_point()
b <- ggplot(new_CreditCard, aes(resid)) + geom_freqpoly(binwidth = 20)
a/b
```

##### Every single increase in depenpents leads to an increase of $11 in average monthly credit card expense. 

```{r echo=FALSE}
m2 <- glm(expenditure ~ dependents, data = new_CreditCard)
summary(m2)
coef(m2)
```

##### Older people are more likely to own a home and stay there for longer periods

```{r echo=FALSE}
ggplot(new_CreditCard, aes(x=age, y=months, colour=owner)) + geom_point() + geom_smooth(method = 'lm', se=FALSE) + labs(title = "Comparing Age and Duration Living in current address", x = "Age", y = "Months")
```

##### Applicants with Derogatory Reports mostly had their applications denied. 

```{r echo=FALSE}
ggplot(new_CreditCard, aes(x=card, y=reports)) + geom_col(fill = 'green') + labs(title = "Reports to Card Approval", x="Card Approval", y="Derogatory Reports")
```

##### Majority of applicants had their cards approved. There is a higher approval rate in the home owners category 

```{r echo=FALSE}
ggplot(new_CreditCard, aes(x=card, fill = owner)) + geom_bar() + labs(title = "Home owner card approval", x="Card Approval", y="Count")
```

##### Only a minority of applicants are self employed. 

```{r echo=FALSE}
ggplot(new_CreditCard, aes(x=card, fill = selfemp)) + geom_bar() + labs(title = "Self Employment Card Approval", x="Card Approval", y="Count")
```

##### Impact of Self Employment on Card Approval. 

Self Employed workers have a slightly lower card approval rate than employed workers

```{r echo=FALSE}
selfemp_prop <- table(new_CreditCard$selfemp, new_CreditCard$card)
#prop.table(selfemp_prop)

ggplot(new_CreditCard, aes(x=selfemp, fill = card)) + geom_bar(position = "fill") + ylab("Proportion") + xlab("Self Employment Status")
```

##### Income doesnt really influence card approval. 

```{r echo=FALSE}
ggplot(new_CreditCard, aes(y=card, x = income)) + geom_boxplot() + labs(title = "Comparing Income to Card Approval", x = "Yearly Income USD10,000", y="Card Approval")
```


## Discussion
The following can be deduced from the analysis:

##### Income
1) Correlation between age and yearly income - The analysis shows that income increases with age. This is more so for employees

##### Expenditure
2) Average monthly credit card expenditure increases with income - From the output of the regression analysis, we can deduce that for every $1,000 increase in yearly income, there is a $45 increase in credit card expenditure
3) Age - No correlation between age and monthly credit card expenditure
3) Dependents - A single increase in the number of dependents leads to an increase of $11 in monthly credit card expenditure

##### Card Approval
4) Derogatory reports reduces chances of card approval
5) Home ownership increases an applicants chances of getting card approval
6) Though self employed applicants have a lower approval rate, the difference is not significant. Hence, we will conclude that self employment status has very little influence on credit card approval
7) Income has very little influence on card approval

##### Prediction
A significant amount of the residuals fall close to 0. Hence, we can conclude that the model is a fairly good prediction model. 


## Conclusion
Credit Card companies should focus on acquiring customers with higher income who have a number of dependents. Analysis shows that these customers will spend more on their credit card and make more revenue for the company.
Card companies should also target home owners; analysis shows that they are more stable(less likely to change location)

### Project Criticism
Further understanding and improvement is required to enhance the level of accuracy of the prediction model.